---
title: "HBCA and PDX Model Label Transfer"
author: "Ryan Davis"
date: "May 4, 2020"
output: html_document
---

## Load R packages

```{r load.packages, echo = FALSE, eval=TRUE}
library(Seurat)
library(dplyr)
library(reshape2)
library(ggplot2)
library(patchwork)
library(cowplot)
library(pheatmap)
library(RColorBrewer)
library(dendsort)
library(viridis)
```

## Combine HBCA Datasets and Subset on Individual

Code to subset the combined HBCA epithelial compartment (basal, luminal 1, luminal 2) into seperate individual seurat objects to uploaod to GitHub. Will not be run after generating each subset.

```{r subset.hbca.dataset, eval = FALSE, echo = TRUE}
#HBCA
hbca.basal<-readRDS("E:/Scripts/Seurat/Seurat3_Objects/hbca_basal.rds")
hbca.basal[["coarse.ident"]]<-rep("basal", nrow(hbca.basal@meta.data))
hbca.luminal1<-readRDS("E:/Scripts/Seurat/Seurat3_Objects/hbca_luminal1.rds")
hbca.luminal1[["coarse.ident"]]<-rep("luminal1", nrow(hbca.luminal1@meta.data))
hbca.luminal2<-readRDS("E:/Scripts/Seurat/Seurat3_Objects/hbca_luminal2.rds")
hbca.luminal2[["coarse.ident"]]<-rep("luminal2", nrow(hbca.luminal2@meta.data))

hbca.combined<-merge(x = hbca.basal,
                     y = c(hbca.luminal1,hbca.luminal2))
hbca.patient.list <- SplitObject(hbca.combined, split.by = "patient_id")

hbca.patients <- names(hbca.patient.list)
for (i in 1:length(hbca.patient.list)){
  assign(paste(hbca.patients[i]), hbca.patient.list[[i]])
}

for (i in 1:length(hbca.patient.list)){
  object<-hbca.patient.list[[i]]
  saveRDS(object, file = paste("C:/Users/Ryan/Documents/pdx-hbca-integration/seurat-objects/",hbca.patients[i],"_seurat3_object.rds",sep = ""))
  print(paste("C:/Users/Ryan/Documents/pdx-hbca-integration/seurat-objects/",hbca.patients[i],"_seurat3_object.rds",sep = ""))
}
```
## Calculate HBCA Integration by Patient

Calculate data integration by patient using log-normalization. Datasets pt8, pt20, pt22, pt25, pt26, pt27, pt28, pt31, pt35 had too few cells, and I removed them from the integration

```{r hbca.dataset.integration, eval = FALSE, echo = TRUE}
#HBCA
load("hbca_patient_names.txt")
hbca.patient.list<-list()
for (i in 1:length(hbca.patients)){
  hbca.patient.list[[i]]<-assign(paste(hbca.patients[i]),
         readRDS(paste("seurat-objects/",hbca.patients[i],"_seurat3_object.rds",sep = "")))
  names(hbca.patient.list)[i]<-paste(hbca.patients[i])
}
#Log Normalize Data
for (i in 1:length(hbca.patient.list)) {
    hbca.patient.list[[i]] <- NormalizeData(hbca.patient.list[[i]], verbose = FALSE, assay = "RNA")
    hbca.patient.list[[i]] <- FindVariableFeatures(hbca.patient.list[[i]], selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE, assay = "RNA")
}
#Remove patients with too few cells
hbca.patient.list$pt8<-NULL
hbca.patient.list$pt20<-NULL
hbca.patient.list$pt22<-NULL
hbca.patient.list$pt25<-NULL
hbca.patient.list$pt26<-NULL
hbca.patient.list$pt27<-NULL
hbca.patient.list$pt28<-NULL
hbca.patient.list$pt31<-NULL
hbca.patient.list$pt35<-NULL

#Find Integration Anchors
hbca.patient.anchors <- FindIntegrationAnchors(object.list = hbca.patient.list, dims = 1:30, k.filter = 100, k.anchor = 2, k.score = 20)
#Integrate Data
hbca.integrated <- IntegrateData(anchorset = hbca.patient.anchors, dims = 1:30)
#Dimentionality Reduction
DefaultAssay(hbca.integrated) <- "integrated"
# Run the standard workflow for visualization and clustering
hbca.integrated <- ScaleData(hbca.integrated, verbose = FALSE)
hbca.integrated <- RunPCA(hbca.integrated, npcs = 30, verbose = FALSE)
hbca.integrated <- RunUMAP(hbca.integrated, reduction = "pca", dims = 1:30)
p1 <- DimPlot(hbca.integrated, reduction = "umap", group.by = "coarse.ident")
p2 <- DimPlot(hbca.integrated, reduction = "umap", group.by = "patient_id", label = FALSE, 
    repel = TRUE)
save(hbca.integrated, file = "E:/Scripts/Seurat/Seurat3_Objects/hbca.integrated.Seurat3.object.Rda")
```

## Load in seurat objects

```{r load.objects, eval = TRUE, echo = TRUE}
#HCI001
load("seurat-objects/hci001.expected.counts.trimmed.seurat3.object.Rda")
#HCI002
load("seurat-objects/hci002.expected.counts.trimmed.seurat3.object.Rda")
#HCI010
load("seurat-objects/hci010.expected.counts.trimmed.seurat3.object.Rda")
#HBCA
load("E:/Scripts/Seurat/Seurat3_Objects/hbca.integrated.Seurat3.object.Rda")
```

## Label Transfer Calculation

```{r label.transfer.prep, eval=TRUE, echo=FALSE}

#Load TPM Seurat Objects to Transfer MetaData
load("seurat-objects/hci001.seurat3.object.Rda")
load("seurat-objects/hci002.seurat3.object.Rda")
load("seurat-objects/hci010.seurat3.object.Rda")

#Reorder HCI001 Meta Data
p01.cc.updated[["cell.names"]]<-rownames(p01.cc.updated@meta.data)
p01.cc.updated@meta.data<-arrange(p01.cc.updated@meta.data, cell.names)
rownames(p01.cc.updated@meta.data)<-as.character(p01.cc.updated$cell.names)

p01.counts[["cell.names"]]<-rownames(p01.counts@meta.data)
p01.counts@meta.data<-arrange(p01.counts@meta.data, cell.names)
rownames(p01.counts@meta.data)<-as.character(p01.counts$cell.names)

#Add HCI001 Meta Data to Counts Seurat Object
p01.counts[["mouse"]]<-p01.cc.updated$mouse
p01.counts[["burden"]]<-p01.cc.updated$burden
p01.counts[["tissue"]]<-p01.cc.updated$tissue
p01.counts[["paper.ident"]]<-p01.cc.updated$paper.ident
p01.counts[["pdx.surgery.type"]]<-p01.cc.updated$pdx.surgery.type
p01.counts[["pdx.surgical.side"]]<-p01.cc.updated$pdx.surgical.side

#Reorder HCI002 Meta Data
p02.updated[["cell.names"]]<-rownames(p02.updated@meta.data)
p02.updated@meta.data<-arrange(p02.updated@meta.data, cell.names)
rownames(p02.updated@meta.data)<-as.character(p02.updated$cell.names)

p02.counts[["cell.names"]]<-rownames(p02.counts@meta.data)
p02.counts@meta.data<-arrange(p02.counts@meta.data, cell.names)
rownames(p02.counts@meta.data)<-as.character(p02.counts$cell.names)

#Add HCI002 Meta Data to Counts Seurat Object
p02.counts[["mouse"]]<-p02.updated$mouse
p02.counts[["burden"]]<-p02.updated$burden
p02.counts[["tissue"]]<-p02.updated$tissue
p02.counts[["paper.ident"]]<-p02.updated$paper.ident
p02.counts[["pdx.surgery.type"]]<-p02.updated$pdx.surgery.type
p02.counts[["pdx.surgical.side"]]<-p02.updated$pdx.surgical.side

#Reorder HCI010 Meta Data
p10.updated[["cell.names"]]<-rownames(p10.updated@meta.data)
p10.updated@meta.data<-arrange(p10.updated@meta.data, cell.names)
rownames(p10.updated@meta.data)<-as.character(p10.updated$cell.names)

p10.counts[["cell.names"]]<-rownames(p10.counts@meta.data)
p10.counts@meta.data<-arrange(p10.counts@meta.data, cell.names)
rownames(p10.counts@meta.data)<-as.character(p10.counts$cell.names)

#Add HCI010 Meta Data to Counts Seurat Object
p10.counts[["mouse"]]<-p10.updated$mouse
p10.counts[["burden"]]<-p10.updated$burden
p10.counts[["tissue"]]<-p10.updated$tissue
p10.counts[["paper.ident"]]<-p10.updated$paper.ident
p10.counts[["pdx.surgery.type"]]<-p10.updated$pdx.surgery.type
p10.counts[["pdx.surgical.side"]]<-p10.updated$pdx.surgical.side
```
```{r label.transfer.calculation, eval=FALSE, echo=TRUE}
options(future.globals.maxSize = (6000*1024^2))
memory.size(max = TRUE)
DefaultAssay(hbca.integrated)<-"integrated"
Idents(hbca.integrated)<-"orig.ident"
hbca.integrated.downsample<-SubsetData(hbca.integrated, max.cells.per.ident = 30000)

hci001.query <- p01.counts
hci001.query<-NormalizeData(hci001.query)
hci001.query<-FindVariableFeatures(hci001.query, selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE)
DefaultAssay(hbca.integrated.downsample)<-"integrated"
hci001.anchors <- FindTransferAnchors(reference = hbca.integrated.downsample, query = hci001.query, 
    dims = 1:30)
predictions <- TransferData(anchorset = hci001.anchors, refdata = hbca.integrated.downsample$coarse.ident, 
    dims = 1:30)
hci001.query <- AddMetaData(hci001.query, metadata = predictions)

hci002.query <- p02.counts
hci002.query<-NormalizeData(hci002.query)
hci002.query<-FindVariableFeatures(hci002.query, selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE)
DefaultAssay(hbca.integrated.downsample)<-"integrated"
hci002.anchors <- FindTransferAnchors(reference = hbca.integrated.downsample, query = hci002.query, 
    dims = 1:30)
predictions <- TransferData(anchorset = hci002.anchors, refdata = hbca.integrated.downsample$coarse.ident, 
    dims = 1:30)
hci002.query <- AddMetaData(hci002.query, metadata = predictions)

hci010.query <- p10.counts
hci010.query<-NormalizeData(hci010.query)
hci010.query<-FindVariableFeatures(hci010.query, selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE)
DefaultAssay(hbca.integrated.downsample)<-"integrated"
hci010.anchors <- FindTransferAnchors(reference = hbca.integrated.downsample, query = hci010.query, 
    dims = 1:30)
predictions <- TransferData(anchorset = hci010.anchors, refdata = hbca.integrated.downsample$coarse.ident, 
    dims = 1:30)
hci010.query <- AddMetaData(hci010.query, metadata = predictions)
```


## Save and Load Label Transfer Objects

```{r label.transfer.save.objects, eval=FALSE, echo=TRUE}
save(hci001.anchors, file = "seurat-objects/HCI001_HBCA_Anchorset_Object.Rda")
save(hci002.anchors, file = "seurat-objects/HCI002_HBCA_Anchorset_Object.Rda")
save(hci010.anchors, file = "seurat-objects/HCI010_HBCA_Anchorset_Object.Rda")
save(hci001.query, file = "seurat-objects/HCI001_HBCA_LabelTransfer_Seurat3_Object.Rda")
save(hci002.query, file = "seurat-objects/HCI002_HBCA_LabelTransfer_Seurat3_Object.Rda")
save(hci010.query, file = "seurat-objects/HCI010_HBCA_LabelTransfer_Seurat3_Object.Rda")
```
```{r label.transfer.load.objects, eval=TRUE, echo=TRUE}
load(file = "seurat-objects/HCI001_HBCA_Anchorset_Object.Rda")
load(file = "seurat-objects/HCI002_HBCA_Anchorset_Object.Rda")
load(file = "seurat-objects/HCI010_HBCA_Anchorset_Object.Rda")
load(file = "seurat-objects/HCI001_HBCA_LabelTransfer_Seurat3_Object.Rda")
load(file = "seurat-objects/HCI002_HBCA_LabelTransfer_Seurat3_Object.Rda")
load(file = "seurat-objects/HCI010_HBCA_LabelTransfer_Seurat3_Object.Rda")
```

##Visualize Label Transfer Identities

```{r label.transfer.visualize.hci001, eval=TRUE, echo=FALSE, fig.height=4, fig.width=7}
hci001.prediction.plot<-FetchData(hci001.query, c("tissue",
                                                  "paper.ident",
                                                  "predicted.id",
                                                  "cell.names",
                                                  "prediction.score.luminal2",
                                                  "prediction.score.luminal1",
                                                  "prediction.score.basal"))
levels(hci001.prediction.plot$tissue)<-c("Metastatic","Tumor")
hci001.prediction.plot.melted<-melt(hci001.prediction.plot)
ggplot(hci001.prediction.plot.melted, aes(value, fill=tissue))+
  geom_histogram()+
  ggtitle("Prediction Score HCI001")+
  theme_classic()+
  facet_wrap(~ variable)

hci001.prediction.heatmap.labels<-hci001.prediction.plot[,1:2]
colnames(hci001.prediction.heatmap.labels)<-c("Status","Paper ID")

colors<-brewer.pal(5, "Dark2")
hci001.prediction.heatmap.colors<-list("Status" = c("Tumor"=colors[1],"Metastatic"=colors[2]),
                                       "Paper ID" = c("A1"=colors[3],"A2"=colors[4],"A3"=colors[5]))

hci001.prediction.heatmap<-t(hci001.prediction.plot[,5:7])
rownames(hci001.prediction.heatmap)<-c("Luminal 2",
                                       "Luminal 1",
                                       "Basal")

sort_hclust <- function(...) as.hclust(dendsort(as.dendrogram(...)))
mat_cluster_cols <- hclust(dist(t((as.matrix(hci001.prediction.heatmap)))))
mat_cluster_cols <- sort_hclust(mat_cluster_cols)
mat_cluster_rows <- sort_hclust(hclust(dist(as.matrix(hci001.prediction.heatmap))))

pheatmap((as.matrix(hci001.prediction.heatmap)),
         scale = "column",
         show_colnames = F,
         cluster_cols = mat_cluster_cols,
         cluster_rows = mat_cluster_rows,
         color = PurpleAndYellow(50),
         annotation_col = hci001.prediction.heatmap.labels,
         annotation_colors = hci001.prediction.heatmap.colors,
         cutree_cols = 1)
```
```{r label.transfer.visualize.hci002, eval=TRUE, echo=FALSE, fig.height=4, fig.width=8}
hci002.prediction.plot<-FetchData(hci002.query, c("tissue",
                                                  "paper.ident",
                                                  "predicted.id",
                                                  "cell.names",
                                                  "prediction.score.luminal2",
                                                  "prediction.score.luminal1",
                                                  "prediction.score.basal"))
levels(hci002.prediction.plot$tissue)<-c("Metastatic","Metastatic","Tumor")
hci002.prediction.plot.melted<-melt(hci002.prediction.plot)
ggplot(hci002.prediction.plot.melted, aes(value, fill=tissue))+
  geom_histogram()+
  ggtitle("Prediction Score hci002")+
  theme_classic()+
  facet_wrap(~ variable)

hci002.prediction.heatmap.labels<-hci002.prediction.plot[,1:2]
colnames(hci002.prediction.heatmap.labels)<-c("Status","Paper ID")

colors<-brewer.pal(7, "Dark2")
hci002.prediction.heatmap.colors<-list("Status" = c("Tumor"=colors[1],"Metastatic"=colors[2]),
                                       "Paper ID" = c("B1"=colors[3],"B2"=colors[4],"B3"=colors[5],"B4"=colors[6],"B5"=colors[7]))

hci002.prediction.heatmap<-t(hci002.prediction.plot[,5:7])
rownames(hci002.prediction.heatmap)<-c("Luminal 2",
                                       "Luminal 1",
                                       "Basal")

sort_hclust <- function(...) as.hclust(dendsort(as.dendrogram(...)))
mat_cluster_cols <- hclust(dist(t((as.matrix(hci002.prediction.heatmap)))))
mat_cluster_cols <- sort_hclust(mat_cluster_cols)
mat_cluster_rows <- sort_hclust(hclust(dist(as.matrix(hci002.prediction.heatmap))))

pheatmap((as.matrix(hci002.prediction.heatmap)),
         scale = "column",
         show_colnames = F,
         cluster_cols = mat_cluster_cols,
         cluster_rows = mat_cluster_rows,
         color = PurpleAndYellow(50),
         annotation_col = hci002.prediction.heatmap.labels,
         annotation_colors = hci002.prediction.heatmap.colors,
         cutree_cols = 1)
```
```{r label.transfer.visualize.hci010, eval=TRUE, echo=FALSE, fig.height=4, fig.width=9}
hci010.prediction.plot<-FetchData(hci010.query, c("tissue",
                                                  "paper.ident",
                                                  "predicted.id",
                                                  "cell.names",
                                                  "prediction.score.luminal2",
                                                  "prediction.score.luminal1",
                                                  "prediction.score.basal"))
levels(hci010.prediction.plot$tissue)<-c("Metastatic","Metastatic","Tumor")
hci010.prediction.plot.melted<-melt(hci010.prediction.plot)

ggplot(hci010.prediction.plot.melted, aes(value, fill=tissue))+
  geom_histogram(binwidth = 0.01)+
  ggtitle("Prediction Score hci010")+
  theme_classic()+
  facet_wrap(~ variable)

hci010.prediction.heatmap.labels<-hci010.prediction.plot[,1:2]
colnames(hci010.prediction.heatmap.labels)<-c("Status","Paper ID")

colors<-brewer.pal(8, "Dark2")
hci010.prediction.heatmap.colors<-list("Status" = c("Tumor"=colors[1],"Metastatic"=colors[2]),
                                       "Paper ID" = c("C1"=colors[3],"C2"=colors[4],"C3"=colors[5],"C4"=colors[6],"C5"=colors[7],"C6"=colors[8]))

hci010.prediction.heatmap<-t(hci010.prediction.plot[,5:7])
rownames(hci010.prediction.heatmap)<-c("Luminal 2",
                                       "Luminal 1",
                                       "Basal")

sort_hclust <- function(...) as.hclust(dendsort(as.dendrogram(...)))
mat_cluster_cols <- hclust(dist(t((as.matrix(hci010.prediction.heatmap)))))
mat_cluster_cols <- sort_hclust(mat_cluster_cols)
mat_cluster_rows <- sort_hclust(hclust(dist(as.matrix(hci010.prediction.heatmap))))

pheatmap((as.matrix(hci010.prediction.heatmap)),
         scale = "column",
         show_colnames = F,
         cluster_cols = mat_cluster_cols,
         cluster_rows = mat_cluster_rows,
         color = PurpleAndYellow(50),
         annotation_col = hci010.prediction.heatmap.labels,
         annotation_colors = hci010.prediction.heatmap.colors)
```
```{r label.transfer.identify.marker.genes.hbca, eval=FALSE, echo=TRUE}
DefaultAssay(hbca.integrated)<-"RNA"
Idents(hbca.integrated)<-"coarse.ident"
hbca.epithelial.markers<-FindAllMarkers(hbca.integrated, only.pos = T)
write.table(hbca.epithelial.markers, file = "marker-genes/hbca.epithelial.markers.txt", sep = "\t")
```

## Visualizing Tranfer Anchor Genes

```{r label.transfer.visualize.transfer.genes.hci001, eval=TRUE, echo=TRUE}
hci001.query<-ScaleData(hci001.query)


hci001.gene.anchors<-hci001.anchors@anchor.features

length(which(hci001.gene.anchors %in% rownames(hci001.query)))
hci001.anchors.plot<-FetchData(hci001.query, c("tissue",
                                                  "paper.ident",
                                                  "predicted.id",
                                                  "cell.names",
                                                  hci001.gene.anchors), slot = "scale.data")
levels(hci001.anchors.plot$tissue)<-c("Metastatic","Tumor")

#Specify Cell Labels (column annotations)
hci001.anchors.heatmap.labels.column<-data.frame(Status=hci001.anchors.plot$tissue)
rownames(hci001.anchors.heatmap.labels.column)<-rownames(hci001.anchors.plot)
#Specify Colors for Annotations
colors<-brewer.pal(5, "Dark2")
hci001.anchors.heatmap.colors.column<-list("Status" = c("Tumor"=colors[1],"Metastatic"=colors[2]),
                                       "gene.type" = c("basal"=colors[3],"luminal1"=colors[4],"luminal2"=colors[5]))
#Sort the dendrograms in order of splits for visualization
hci001.anchors.heatmap<-t(hci001.anchors.plot[,5:ncol(hci001.anchors.plot)])
sort_hclust <- function(...) as.hclust(dendsort(as.dendrogram(...)))
mat_cluster_cols <- hclust(dist(t((as.matrix(hci001.anchors.heatmap)))))
mat_cluster_cols <- sort_hclust(mat_cluster_cols)
mat_cluster_rows <- sort_hclust(hclust(dist(as.matrix(hci001.anchors.heatmap))))
#Set the breaks for the color scale of the heatmap
quantile_breaks <- function(xs, n = 10) {
  breaks <- quantile(xs, probs = seq(0, 1, length.out = n))
  breaks[!duplicated(breaks)]
}
mat_breaks <- quantile_breaks(hci001.anchors.heatmap, n = 11)
#Add labels for genes and colors
hbca.epithelial.markers<-read.table(file = "marker-genes/hbca.epithelial.markers.txt", sep = "\t")
hci001.anchors.heatmap.labels.row<-data.frame(gene.type=hbca.epithelial.markers$cluster)
rownames(hci001.anchors.heatmap.labels.row)<-rownames(hbca.epithelial.markers)

#Run the heatmap with all genes
pheatmap((as.matrix(hci001.anchors.heatmap)),
         scale = "none",
         show_colnames = F,
         breaks = mat_breaks,
         cluster_cols = mat_cluster_cols,
         cluster_rows = mat_cluster_rows,
         color = PurpleAndYellow((length(mat_breaks) - 1)),
         annotation_col = hci001.anchors.heatmap.labels.column,
         annotation_colors = hci001.anchors.heatmap.colors.column,
         annotation_row = hci001.anchors.heatmap.labels.row,
         annotation_names_row = F,
         show_rownames = FALSE,
         main = "All Transfer Anchor Genes (n=1400)")

#Sort the dendrograms in order of splits for visualization
hci001.anchors.heatmap.reduced<-t(hci001.anchors.plot[,which(colnames(hci001.anchors.plot) %in% hbca.epithelial.markers$gene)])
sort_hclust <- function(...) as.hclust(dendsort(as.dendrogram(...)))
mat_cluster_cols <- hclust(dist(t((as.matrix(hci001.anchors.heatmap.reduced)))))
mat_cluster_cols <- sort_hclust(mat_cluster_cols)
mat_cluster_rows <- sort_hclust(hclust(dist(as.matrix(hci001.anchors.heatmap.reduced))))
#Set the breaks for the color scale of the heatmap.reduced
quantile_breaks <- function(xs, n = 10) {
  breaks <- quantile(xs, probs = seq(0, 1, length.out = n))
  breaks[!duplicated(breaks)]
}
mat_breaks <- quantile_breaks(hci001.anchors.heatmap.reduced, n = 11)
#Add labels for genes and colors
hbca.epithelial.markers<-read.table(file = "marker-genes/hbca.epithelial.markers.txt", sep = "\t")
hci001.anchors.heatmap.reduced.labels.row<-data.frame(gene.type=hbca.epithelial.markers$cluster)
rownames(hci001.anchors.heatmap.reduced.labels.row)<-rownames(hbca.epithelial.markers)

#Run the heatmap.reduced with all genes
pheatmap((as.matrix(hci001.anchors.heatmap.reduced)),
         scale = "none",
         show_colnames = F,
         breaks = mat_breaks,
         cluster_cols = mat_cluster_cols,
         cluster_rows = mat_cluster_rows,
         color = PurpleAndYellow((length(mat_breaks) - 1)),
         annotation_col = hci001.anchors.heatmap.labels.column,
         annotation_colors = hci001.anchors.heatmap.colors.column,
         annotation_row = hci001.anchors.heatmap.labels.row,
         annotation_names_row = F,
         show_rownames = FALSE,
         main = "Transfer Anchor Genes Subset by Markers (n=615)")
```
```{r label.transfer.visualize.transfer.genes.hci002, eval=TRUE, echo=TRUE}
hci002.query<-NormalizeData(hci002.query)
hci002.query<-ScaleData(hci002.query)


hci002.gene.anchors<-hci002.anchors@anchor.features

length(which(hci002.gene.anchors %in% rownames(hci002.query)))
hci002.anchors.plot<-FetchData(hci002.query, c("tissue",
                                                  "paper.ident",
                                                  "predicted.id",
                                                  "cell.names",
                                                  hci002.gene.anchors), slot = "scale.data")
levels(hci002.anchors.plot$tissue)<-c("Metastatic","Metastatic","Tumor")

#Specify Cell Labels (column annotations)
hci002.anchors.heatmap.labels.column<-data.frame(Status=hci002.anchors.plot$tissue)
rownames(hci002.anchors.heatmap.labels.column)<-rownames(hci002.anchors.plot)
#Specify Colors for Annotations
colors<-brewer.pal(5, "Dark2")
hci002.anchors.heatmap.colors.column<-list("Status" = c("Tumor"=colors[1],"Metastatic"=colors[2]),
                                       "gene.type" = c("basal"=colors[3],"luminal1"=colors[4],"luminal2"=colors[5]))
#Sort the dendrograms in order of splits for visualization
hci002.anchors.heatmap<-t(hci002.anchors.plot[,5:ncol(hci002.anchors.plot)])
sort_hclust <- function(...) as.hclust(dendsort(as.dendrogram(...)))
mat_cluster_cols <- hclust(dist(t((as.matrix(hci002.anchors.heatmap)))))
mat_cluster_cols <- sort_hclust(mat_cluster_cols)
mat_cluster_rows <- sort_hclust(hclust(dist(as.matrix(hci002.anchors.heatmap))))
#Set the breaks for the color scale of the heatmap
quantile_breaks <- function(xs, n = 10) {
  breaks <- quantile(xs, probs = seq(0, 1, length.out = n))
  breaks[!duplicated(breaks)]
}
mat_breaks <- quantile_breaks(hci002.anchors.heatmap, n = 11)
#Add labels for genes and colors
hbca.epithelial.markers<-read.table(file = "marker-genes/hbca.epithelial.markers.txt", sep = "\t")
hci002.anchors.heatmap.labels.row<-data.frame(gene.type=hbca.epithelial.markers$cluster)
rownames(hci002.anchors.heatmap.labels.row)<-rownames(hbca.epithelial.markers)

#Run the heatmap with all genes
pheatmap((as.matrix(hci002.anchors.heatmap)),
         scale = "none",
         show_colnames = F,
         breaks = mat_breaks,
         cluster_cols = mat_cluster_cols,
         cluster_rows = mat_cluster_rows,
         color = PurpleAndYellow((length(mat_breaks) - 1)),
         annotation_col = hci002.anchors.heatmap.labels.column,
         annotation_colors = hci002.anchors.heatmap.colors.column,
         annotation_row = hci002.anchors.heatmap.labels.row,
         annotation_names_row = F,
         show_rownames = FALSE,
         main = "All Transfer Anchor Genes (n=1346)")

#Sort the dendrograms in order of splits for visualization
hci002.anchors.heatmap.reduced<-t(hci002.anchors.plot[,which(colnames(hci002.anchors.plot) %in% hbca.epithelial.markers$gene)])
sort_hclust <- function(...) as.hclust(dendsort(as.dendrogram(...)))
mat_cluster_cols <- hclust(dist(t((as.matrix(hci002.anchors.heatmap.reduced)))))
mat_cluster_cols <- sort_hclust(mat_cluster_cols)
mat_cluster_rows <- sort_hclust(hclust(dist(as.matrix(hci002.anchors.heatmap.reduced))))
#Set the breaks for the color scale of the heatmap.reduced
quantile_breaks <- function(xs, n = 10) {
  breaks <- quantile(xs, probs = seq(0, 1, length.out = n))
  breaks[!duplicated(breaks)]
}
mat_breaks <- quantile_breaks(hci002.anchors.heatmap.reduced, n = 11)
#Add labels for genes and colors
hbca.epithelial.markers<-read.table(file = "marker-genes/hbca.epithelial.markers.txt", sep = "\t")
hci002.anchors.heatmap.reduced.labels.row<-data.frame(gene.type=hbca.epithelial.markers$cluster)
rownames(hci002.anchors.heatmap.reduced.labels.row)<-rownames(hbca.epithelial.markers)

#Run the heatmap.reduced with all genes
pheatmap((as.matrix(hci002.anchors.heatmap.reduced)),
         scale = "none",
         show_colnames = F,
         breaks = mat_breaks,
         cluster_cols = mat_cluster_cols,
         cluster_rows = mat_cluster_rows,
         color = PurpleAndYellow((length(mat_breaks) - 1)),
         annotation_col = hci002.anchors.heatmap.labels.column,
         annotation_colors = hci002.anchors.heatmap.colors.column,
         annotation_row = hci002.anchors.heatmap.labels.row,
         annotation_names_row = F,
         show_rownames = FALSE,
         main = "Transfer Anchor Genes Subset by Markers (n=615)")
```
```{r label.transfer.visualize.transfer.genes.hci010, eval=TRUE, echo=TRUE}
hci010.query<-NormalizeData(hci010.query)
hci010.query<-ScaleData(hci010.query)


hci010.gene.anchors<-hci010.anchors@anchor.features

length(which(hci010.gene.anchors %in% rownames(hci010.query)))
hci010.anchors.plot<-FetchData(hci010.query, c("tissue",
                                                  "paper.ident",
                                                  "predicted.id",
                                                  "cell.names",
                                                  hci010.gene.anchors), slot = "scale.data")
levels(hci010.anchors.plot$tissue)<-c("Metastatic","Metastatic","Tumor")

#Specify Cell Labels (column annotations)
hci010.anchors.heatmap.labels.column<-data.frame(Status=hci010.anchors.plot$tissue)
rownames(hci010.anchors.heatmap.labels.column)<-rownames(hci010.anchors.plot)
#Specify Colors for Annotations
colors<-brewer.pal(5, "Dark2")
hci010.anchors.heatmap.colors.column<-list("Status" = c("Tumor"=colors[1],"Metastatic"=colors[2]),
                                       "gene.type" = c("basal"=colors[3],"luminal1"=colors[4],"luminal2"=colors[5]))
#Sort the dendrograms in order of splits for visualization
hci010.anchors.heatmap<-t(hci010.anchors.plot[,5:ncol(hci010.anchors.plot)])
sort_hclust <- function(...) as.hclust(dendsort(as.dendrogram(...)))
mat_cluster_cols <- hclust(dist(t((as.matrix(hci010.anchors.heatmap)))))
mat_cluster_cols <- sort_hclust(mat_cluster_cols)
mat_cluster_rows <- sort_hclust(hclust(dist(as.matrix(hci010.anchors.heatmap))))
#Set the breaks for the color scale of the heatmap
quantile_breaks <- function(xs, n = 10) {
  breaks <- quantile(xs, probs = seq(0, 1, length.out = n))
  breaks[!duplicated(breaks)]
}
mat_breaks <- quantile_breaks(hci010.anchors.heatmap, n = 11)
#Add labels for genes and colors
hbca.epithelial.markers<-read.table(file = "marker-genes/hbca.epithelial.markers.txt", sep = "\t")
hci010.anchors.heatmap.labels.row<-data.frame(gene.type=hbca.epithelial.markers$cluster)
rownames(hci010.anchors.heatmap.labels.row)<-rownames(hbca.epithelial.markers)

#Run the heatmap with all genes
pheatmap((as.matrix(hci010.anchors.heatmap)),
         scale = "none",
         show_colnames = F,
         breaks = mat_breaks,
         cluster_cols = mat_cluster_cols,
         cluster_rows = mat_cluster_rows,
         color = PurpleAndYellow((length(mat_breaks) - 1)),
         annotation_col = hci010.anchors.heatmap.labels.column,
         annotation_colors = hci010.anchors.heatmap.colors.column,
         annotation_row = hci010.anchors.heatmap.labels.row,
         annotation_names_row = F,
         show_rownames = FALSE,
         main = "All Transfer Anchor Genes (n=1546)")

#Sort the dendrograms in order of splits for visualization
hci010.anchors.heatmap.reduced<-t(hci010.anchors.plot[,which(colnames(hci010.anchors.plot) %in% hbca.epithelial.markers$gene)])
sort_hclust <- function(...) as.hclust(dendsort(as.dendrogram(...)))
mat_cluster_cols <- hclust(dist(t((as.matrix(hci010.anchors.heatmap.reduced)))))
mat_cluster_cols <- sort_hclust(mat_cluster_cols)
mat_cluster_rows <- sort_hclust(hclust(dist(as.matrix(hci010.anchors.heatmap.reduced))))
#Set the breaks for the color scale of the heatmap.reduced
quantile_breaks <- function(xs, n = 10) {
  breaks <- quantile(xs, probs = seq(0, 1, length.out = n))
  breaks[!duplicated(breaks)]
}
mat_breaks <- quantile_breaks(hci010.anchors.heatmap.reduced, n = 11)
#Add labels for genes and colors
hbca.epithelial.markers<-read.table(file = "marker-genes/hbca.epithelial.markers.txt", sep = "\t")
hci010.anchors.heatmap.reduced.labels.row<-data.frame(gene.type=hbca.epithelial.markers$cluster)
rownames(hci010.anchors.heatmap.reduced.labels.row)<-rownames(hbca.epithelial.markers)

#Run the heatmap.reduced with all genes
pheatmap((as.matrix(hci010.anchors.heatmap.reduced)),
         scale = "none",
         show_colnames = F,
         breaks = mat_breaks,
         cluster_cols = mat_cluster_cols,
         cluster_rows = mat_cluster_rows,
         color = PurpleAndYellow((length(mat_breaks) - 1)),
         annotation_col = hci010.anchors.heatmap.labels.column,
         annotation_colors = hci010.anchors.heatmap.colors.column,
         annotation_row = hci010.anchors.heatmap.labels.row,
         annotation_names_row = F,
         show_rownames = FALSE,
         main = "Transfer Anchor Genes Subset by Markers (n=654)")
```

## Identifying Cellular Anchors

```{r label.transfer.cell.anchors.hci001, eval=TRUE, echo=FALSE}
reference.anchors<-hci001.anchors@reference.cells[hci001.anchors@anchors[,1]]
#Remove the "_reference" from the cell names
reference.anchors<-gsub("_reference","",reference.anchors)
hbca.reference.object<-SubsetData(hbca.integrated, cells = reference.anchors)
table(hbca.reference.object@meta.data$coarse.ident)

query.anchors<-hci001.anchors@query.cells[hci001.anchors@anchors[,2]]
#Remove the "_query" from the cell names
query.anchors<-gsub("_query","",query.anchors)
hci001.query.object<-SubsetData(hci001.query, cells = query.anchors)
table(hci001.query.object@meta.data$tissue)

```